Graph Neural Networks For Multi-Image Matching

### 2019-09-23 latex/iclr2020/main.tex
Added graffiti text and a few minor typos

\begin{abstract}
    Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to acquire.
    In addition, geometric consistency losses can be added at training time, even if the information is not available in the test set, unlike previous approaches which optimize cycle consistency directly.
    To the best of our knowledge, no other works have used learning for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section 3.3).
\cite{pachauri2013solving} and \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the low Gaussian noise assumption is not realistic.
\cite{zhou2015multi} and \cite{wang2017multi} use most sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
\cite{leonardos2016distributed} implement a distributed optimization scheme to solve for cycle consistency.
As an alternative to optimization based techniques, \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
To the best of our knowledge, we are the first to use neural networks for multi-image matching.

@@

\subsection{Graph Neural Networks}
Graph neural networks have recieved more attention recently e.g. \cite{bronstein2017geometric, bruna2013spectral, defferrard2016convolutional, kipf2017semi, scarselli2009graph, gama2018mimo, gama2018convolutional}; and \cite{battaglia2018relational}.
Classical so-called Spectral methods used the eigenvectors of the graph Laplacian to compute convolutions as in \cite{bruna2013spectral}, but requires an a-priori known graph structure. 
Non-spectral methods do not require a-priori knowledge, as seen in \cite{bronstein2017geometric, kipf2017semi, scarselli2009graph}; and \cite{gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
\cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in \cite{battaglia2018relational}.

@@

\begin{figure}[t]
\begin{subfigure}[b]{.55\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{figures-GeometricConsistency-v2.pdf}
  \caption{Illustrated here is an example of how the geometric loss is computed for one feature.}
  \label{fig:1b}
  \label{fig:geoconsist}
\end{subfigure}
\begin{subfigure}[b]{.30\textwidth}
  \centering
  % \includegraphics[width=0.65\linewidth]{figures-GeometricLoss.pdf}
  \includegraphics[height=180px]{figures-GeometricLoss.pdf}
  \caption{}
  \label{fig:3b}
  \label{fig:geomloss}
\end{subfigure}
\caption{
  \textbf{(a)} Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
  The epipolar line is the line of projection of the feature on the first image, projected onto to the second.
  The distance to this line on the second image indicates how likely that point is to correspond geometrically to the original feature.
  There can be false positives along the projected line, as shown by the square feature in the figure, but other points will be eliminated, such as the hexagonal feature.
  Best viewed in color.
  \textbf{(b)} Training curves with and without Geometric Training loss, described in \ref{eq:geom_cost2}.
  The geometric training loss improves testing performance.
  While the training is higher due to the additional loss terms, the ground truth L1 error is substantially better with the Geometric Loss.
  Best viewed in color.
}
\label{fig:1}
\end{figure}

@@

Let $\mM$ be the noiseless set of matches between our features, with $\mM_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then for all $i, j, k$:
\begin{equation}
\mM_{ij} = \mM_{ik} \mM_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency (\cite{zhou2015multi, wang2017multi, leonardos2016distributed}).
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mX_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mM_{ij} = \mX_{i}\mX_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

This reduces the number of our constraints from $O(n^3)$ to $O(n^2)$.
We try to learn vertex embeddings $\mF_V$ to approximate $\mX$ - in other words the final embedding should be an encoding of the universe of features.
As we do not have the ground truth matches $\mM$, we approximate it using the noisy adjacency matrix $\mA$ of our correspondence graph. Thus our loss would be 
\begin{equation}
\mathcal{L}(\mA, \mF_V) = \mathcal{D}(\mA, \mF_V \mF_V^\top)
\end{equation}
Here $\mathcal{D}$ could be an $L_2$ loss, $L_1$ loss, or many others. In this work, we use the $L_1$ loss. 
Note that because of this formulation, we can determine our embeddings only up to a rotation, as
$ \mE R(\mE R)^\top
= \mE RR^\top \mE^\top
= \mE\mE^\top
\label{eq:rotinvar} $
Thus when visualizing embeddings, we rotate them to make them more interpretable (see figure \ref{fig:embeddingsviz}).

@@

One of the main advantages of this approach over more traditional optimization based approaches is the ability to add geometric consistency information into the loss at training time, even if it is not available at test time.
The simplest way to add geometric consistency losses, and the approach we use here, is to use the epipolar constraint.
The epipolar constraint describes how the positions of features in different images corresponding to the same point should be related.
An illustration of this is provided in figure \ref{fig:geoconsist}, showing how this loss can help reject erroneous points.
Given a relative pose $(R_{ij}, T_{ij})$ between two cameras $i$ and $j$  (transforms $j$ to $i$) the epipolar on corresponding feature locations $X_i$ and $X_j$: $X_{i}^\top \cross{T_{ij}}R_{ij} X_{j} = 0$.
In this work we use the two pose epipolar constraint (\cite{tron2014quotient}):
\begin{equation}
X_{i}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j} = 0
\label{eq:essential_constraint}
\end{equation}
% Where the $(R_k, T_k)$ are the poses of cameras $i$ and $j$ respectively.
The constraint assumes that the $X_k$ are calibrated (i.e. the camera intrinsics are known). 
Given our vertex embeddings matrix $\mF_V$, we can formulate a loss between all cameras $i$ and $j$:
\begin{align}
\mathcal{L}_{ij,geom}(\mF_V) = 
&\sum_{v \in \mV(i),u \in \mathcal{V}(j)} (\vf_v \cdot \vf_u) \left|X_{v}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{c_j} X_{u}\right|
\label{eq:geom_cost}
\end{align}
With $\mathcal{V}(i)$ being the vertices associated with camera $i$.
For our purposes, since we use low rank embeddings $\mE_{i}$, $\mE_{j}$, the loss would read (where $c(k)$ is the appropriate camera for point index $k$):
\begin{align} 
\mathcal{L}_{geom}(\mF_V)
=&\; \mathrm{tr}(\mG^\top \mF_V\mF_V^\top) = \sum_{k,l} (\mF_V)_{k} \cdot (\mF_V)_{l} (\mG)_{kl} \label{eq:geom_cost2} \\
(\mG)_{kl} =&\; \left|X_{k}^\top R_{c(k)}^\top \cross{T_{c(l)} - T_{c(k)}}R_{c(l)} X_{l}\right| \nonumber
\end{align}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{figures-error_lines.pdf}
% \includegraphics[height=200px]{figures-error_lines.pdf}
\label{fig:3}
\label{fig:errorlines}
\caption{
  Plot of the losses of the baselines at different iteration numbers.
  The line shows the mean of the graph while the translucent coloring shows the $25^{th}$ to $75^{th}$ percentiles.
  The ROC AUC curves remain fairly consistent while the L1 loss goes noticibly down after more iterations.
  Our method compares to 35-45 iterations of MatchALS, while only having 16 layers and 8 message passes.
  PGDDS performs better than us in $L_1$ but we perform similarly in the ROC AUC metric.
}
\label{fig:3}
\end{figure}

@@

\subsection{Rome 16K Graph Dataset}
We  use the Rome16K dataset (\cite{li2010location}) to test our algorithm in real world settings.
Rome16K consists of 16 thousand images of various historical sites in Rome extracted from Flickr, along with the 3D structure of the sites provided by bundle adjustment.
While not a standard dataset to test cycle consistency, other datasets had insufficient data to train a network on.
Rome16K is typically used to test bundle adjustment methods.
Therefore, to use our method, we extract image triplets and quadruplets with overlap of 80 points or more to test our algorithm, with the points established as corresponding in the given bundle adjustment output.
For the initial embedding we use the original 128 dimensional SIFT descriptors, normalized to have unit $L_2$ norm, the calibrated x-y position, the orientation, and log scale of the SIFT feature.
To construct the graph, we take each feature as a vertex and create edges to the 5 nearest SIFT descriptors for the other images.

@@

>>>>

\begin{abstract}
    Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to acquire.
    In addition, geometric consistency losses can be added at training time, even if the information is not available in the test set, unlike previous approaches which optimize cycle consistency directly.
    To the best of our knowledge, no other works have used graph neural networks for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section 3.3).
\cite{pachauri2013solving} and \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the low Gaussian noise assumption is not realistic.
\cite{zhou2015multi} and \cite{wang2017multi} use most sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
\cite{leonardos2016distributed} implement a distributed optimization scheme to solve for cycle consistency.
As an alternative to optimization based techniques, \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
Moving away from feature matching, \cite{zach2010disambiguating} uses cycle consistency like constraints on pose graphs quite effectively.
To the best of our knowledge, we are the first to use graph neural networks for multi-image matching.

@@

\subsection{Graph Neural Networks}
Graph neural networks have received more attention recently e.g. \cite{bronstein2017geometric, bruna2013spectral, defferrard2016convolutional, kipf2017semi, scarselli2009graph, gama2018mimo, gama2018convolutional}; and \cite{battaglia2018relational}.
Classical so-called Spectral methods used the eigenvectors of the graph Laplacian to compute convolutions as in \cite{bruna2013spectral}, but requires an a-priori known graph structure. 
Non-spectral methods do not require a-priori knowledge, as seen in \cite{bronstein2017geometric, kipf2017semi, scarselli2009graph}; and \cite{gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
\cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in \cite{battaglia2018relational}.

@@

\begin{figure}[t]
\begin{subfigure}[b]{.55\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{figures-GeometricConsistency-v2.pdf}
  \caption{Illustrated here is an example of how the geometric loss is computed for one feature.}
  \label{fig:1b}
  \label{fig:geoconsist}
\end{subfigure}
\begin{subfigure}[b]{.30\textwidth}
  \centering
  % \includegraphics[width=0.65\linewidth]{figures-GeometricLoss.pdf}
  \includegraphics[height=180px]{figures-GeometricLoss.pdf}
  \caption{}
  \label{fig:3b}
  \label{fig:geomloss}
\end{subfigure}
\caption{
  \textbf{(a)} Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
  The epipolar line is the line of projection of the feature on the first image, projected onto to the second.
  The distance to this line on the second image indicates how likely that point is to correspond geometrically to the original feature.
  There can be false positives along the projected line, as shown by the square feature in the figure, but other points will be eliminated, such as the hexagonal feature.
  \textbf{(b)} Training curves with and without Geometric Training loss, described in \ref{eq:geom_cost2}.
  The geometric training loss improves testing performance.
  While the training is higher due to the additional loss terms, the ground truth L1 error is substantially better with the Geometric Loss.
  Best viewed in color.
}
\label{fig:1}
\end{figure}

@@

Let $\mM$ be the noiseless set of matches between our features, with $\mM_{ij}$ being the partial permutation representing the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then for all $i, j, k$:
\begin{equation}
\mM_{ij} = \mM_{ik} \mM_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency (\cite{zhou2015multi, wang2017multi, leonardos2016distributed}).
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mX_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mM_{ij} = \mX_{i}\mX_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

This reduces the number of our constraints from $O(n^3)$ to $O(n^2)$.
We try to learn vertex embeddings $\mF_V$ to approximate $\mX$ - in other words the final embedding should be an encoding of the universe of features.
As we do not have the ground truth matches $\mM$, we approximate it using the noisy adjacency matrix $\mA$ of our correspondence graph. Thus our loss would be 
\begin{equation}
\mathcal{L}(\mA, \mF_V) = \mathcal{D}(\mA, \mF_V \mF_V^\top)
\end{equation}
Here $\mathcal{D}$ could be an $L_2$ loss, $L_1$ loss, or many others. In this work, we use the $L_1$ loss. 
Note that because of this formulation, we can determine our embeddings only up to a rotation, as
$ \mF_V R(\mF_V R)^\top
= \mF_V RR^\top \mF_V^\top
= \mF_V\mF_V^\top
\label{eq:rotinvar} $
Thus when visualizing embeddings, we rotate them to make them more interpretable (see figure \ref{fig:embeddingsviz}).

@@

One of the main advantages of this approach over more traditional optimization based approaches is the ability to add geometric consistency information into the loss at training time, even if it is not available at test time.
The simplest way to add geometric consistency losses, and the approach we use here, is to use the epipolar constraint.
The epipolar constraint describes how the positions of features in different images corresponding to the same point should be related.
An illustration of this is provided in figure \ref{fig:geoconsist}, showing how this loss can help reject erroneous points.
Given a relative pose $(R_{ij}, T_{ij})$ between two cameras $i$ and $j$  (transforms $j$ to $i$) the epipolar on corresponding feature locations $X_i$ and $X_j$: $X_{i}^\top \cross{T_{ij}}R_{ij} X_{j} = 0$.
In this work we use the two pose epipolar constraint (\cite{tron2014quotient}):
\begin{equation}
X_{i}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j} = 0
\label{eq:essential_constraint}
\end{equation}
% Where the $(R_k, T_k)$ are the poses of cameras $i$ and $j$ respectively.
The constraint assumes that the $X_k$ are calibrated (i.e. the camera intrinsics are known). 
Given our vertex embeddings matrix $\mF_V$, we can formulate a loss between all cameras $i$ and $j$:
\begin{align}
\mathcal{L}_{ij,geom}(\mF_V) = 
&\sum_{v \in \mV(i),u \in \mathcal{V}(j)} (\vf_v \cdot \vf_u) \left|X_{v}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{c_j} X_{u}\right|
\label{eq:geom_cost}
\end{align}
With $\mathcal{V}(i)$ being the vertices associated with camera $i$.
For our purposes, since we use low rank embeddings $\mF_{V, i}$, $\mF_{V, j}$, the loss would read (where $c(k)$ is the appropriate camera for point index $k$):
\begin{align} 
\mathcal{L}_{geom}(\mF_V)
=&\; \mathrm{tr}(\mG^\top \mF_V\mF_V^\top) = \sum_{k,l} (\mF_V)_{k} \cdot (\mF_V)_{l} (\mG)_{kl} \label{eq:geom_cost2} \\
(\mG)_{kl} =&\; \left|X_{k}^\top R_{c(k)}^\top \cross{T_{c(l)} - T_{c(k)}}R_{c(l)} X_{l}\right| \nonumber
\end{align}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{figures-error_lines.pdf}
% \includegraphics[height=200px]{figures-error_lines.pdf}
\caption{
  Plot of the losses of the baselines at different iteration numbers.
  The line shows the mean of the graph while the translucent coloring shows the $25^{th}$ to $75^{th}$ percentiles.
  The ROC AUC curves remain fairly consistent while the L1 loss goes noticibly down after more iterations.
  Our method compares to 35-45 iterations of MatchALS, while only having 16 layers and 8 message passes.
  PGDDS performs better than us in $L_1$ but we perform similarly in the ROC AUC metric.
  These results still hold even when we change domains to the Graffiti dataset (see \ref{sec:graffiti}).
}
\label{fig:3}
\label{fig:errorlines}
\end{figure}

@@

\subsection{Rome 16K Graph Dataset}
We  use the Rome16K dataset (\cite{li2010location}) to test our algorithm in real world settings.
Rome16K consists of 16 thousand images of various historical sites in Rome extracted from Flickr, along with the 3D structure of the sites provided by bundle adjustment.
While not a standard dataset to test cycle consistency, other datasets had insufficient data to train a network on.
Rome16K is typically used to test bundle adjustment methods.
Therefore, to use our method, we extract 6-tuples and 10-tuples of images with overlap of 80 points or more to test our algorithm, with the points established as corresponding in the given bundle adjustment output.
For the initial embedding we use the original 128 dimensional SIFT descriptors, normalized to have unit $L_2$ norm, the calibrated x-y position, the orientation, and log scale of the SIFT feature.
To construct the graph, we take each feature as a vertex and create edges to the 5 nearest SIFT descriptors for the other images.

@@

\subsubsection{Graffiti Dataset} \label{sec:graffiti}
We also run our trained model on the Graffiti Dataset (formatted to be able to be input to our model properly). The results are shown in \ref{fig:errorlines} in the rightmost figure. As the graffiti dataset is very small (only 6 views total), we were not able to train on it. We randomly permute the intra-image order of the features to add some variance - by design the GNN outputs the same result each time, while the optimization methods have a very small amount of variance. The transferred results of Graffiti are similar to the test error of Rome16K - smaller $L_1$ error and comparable ROC error. This shows that the GNNs trained in Rome16K generalize similarly to the optimization based methods.
